The effects of binarity on planet occurrence rates measured by transit surveys

### 2017-11-14 paper/introduction.tex
paper: improved clarity; corrected eqn error

A group of astronomers wants to measure the mean number of planets of a 
certain type per star of a certain type.
Ignoring stellar multiplicity, they perform a signal-to-noise limited transit 
survey and detect $N_{\rm det}$ signals that seem to be from the desired 
class of planet.
They count the apparent number of selected stars, $N_\star$.
Assuming their survey is limited by Poisson noise, our astronomers compute 
their detection efficiency $p_{\rm det}$ as a function of planetary and 
stellar properties, in order to know which of the selected stars were
searchable.
Correcting for the geometric transit probability $p_{\rm tra}$, they report an 
apparent occurrence rate $\Lambda_{\rm a}$:
\begin{equation}
\Lambda_{\rm a} = \frac{N_{\rm det}}{N_\star} \times \frac{1}{p_{\rm tra}}
								  \times \frac{1}{p_{\rm det}}.
\label{eq:wouldnt_it_be_nice}
\end{equation}

@@

To gain intuition for the many observational biases at play,
we consider the following idealized transit surveys:
\begin{itemize}
    \item Model \#1: fixed stars, fixed planets, twin binaries;
    \item Model \#2: fixed planets and primaries, varying secondaries;
    \item Model \#3: fixed primaries, varying planets and secondaries.
\end{itemize}
We define our terminology in Sec.~\ref{sec:definitions}, and present
our transit survey models in Secs.~\ref{sec:model_1}-\ref{sec:model_3}, where 
subsequent subsections correspond to the models above.
We interpret these calculations throughout, and in 
Sec.~\ref{sec:discussion} discuss their relevance to topical questions in 
the interpretation of transit survey occurrence rates.
In particular, we mention the ``hot Jupiter rate discrepancy'', the relevance 
towards measurements of $\eta_\oplus$, and the significance for the dearth of 
planets recently discovered by Fulton et al. (2017).
We conclude in Sec.~\ref{sec:conclusion}

>>>>

A group of astronomers wants to measure the mean number of planets of a 
certain type per star of a certain type.
Ignoring stellar multiplicity, they perform a signal-to-noise limited transit 
survey and detect $N_{\rm det}$ signals that seem to be from the desired 
class of planet.
They count the apparent number of selected stars, $N_\star$.
Assuming their survey is limited by Poisson noise, the group computes
their detection efficiency $p_{\rm det}$ as a function of planetary and 
stellar properties, in order to know which of the selected stars were
searchable.
Correcting for the geometric transit probability $p_{\rm tra}$, they report an 
apparent occurrence rate $\Lambda_{\rm a}$:
\begin{equation}
\Lambda_{\rm a} = \frac{N_{\rm det}}{N_\star} \times \frac{1}{p_{\rm tra}}
								  \times \frac{1}{p_{\rm det}}.
\label{eq:wouldnt_it_be_nice}
\end{equation}

@@

To gain intuition for the many observational biases at play,
we consider a set of idealized transit surveys:
\begin{itemize}
    \item Model \#1: fixed stars, fixed planets, twin binaries;
    \item Model \#2: fixed planets and primaries, varying secondaries;
    \item Model \#3: fixed primaries, varying planets and secondaries.
\end{itemize}
We define our terminology in Sec.~\ref{sec:definitions}, and present
our transit survey models in Secs.~\ref{sec:model_1}-\ref{sec:model_3}, where 
each subsection corresponds to a model listed above.
We interpret these calculations throughout, and in 
Sec.~\ref{sec:discussion} connect them to topical questions in 
the interpretation of transit survey occurrence rates.
In particular, we mention the ``hot Jupiter rate discrepancy'', the relevance 
towards measurements of $\eta_\oplus$, and the significance for the dearth of 
planets recently discovered by Fulton et al. (2017).
We conclude in Sec.~\ref{sec:conclusion}

### 2017-11-14 paper/main.tex
paper: improved clarity; corrected eqn error

This work aims to clarify the biases that stellar binarity introduces
to occurrence rates inferred from transit surveys.
In general, stellar multiplicity leads to diluted planetary 
radii, overestimated detection efficiencies, and an undercounted number of 
selected stars (but possibly an overcounted number of searched stars).
These effects skew occurrence rate measurements in different directions, and 
we develop simple models that allow us to understand the crucial effects.
For a model in which all stellar systems are either single or twin binaries, 
and all planets are identical, we find that ignoring binarity leads to an 
underestimate of the occurrence at the true radius by a multiplicative factor 
of $0.8$ (assuming a 10\% twin binary fraction).
Using more realistic models for the stellar population and planetary 
radii, we find that ignoring binarity leads to a $10$\,-\,$30\%$ 
underestimate of the number of planets per star, depending on the true radius 
distribution.
For the occurrence of Earth-sized planets, in our most realistic model the 
underestimate is by $\approx 15\%$~--~at present far smaller than 
systematic uncertainties on real $\eta_\oplus$ measurements.
For hot Jupiters, we find that the inferred occurrence rate is 
$\approx 1.3\times$ smaller than the true rate around single stars.
We suggest that this latter effect contributes to the discrepant hot Jupiter 
rates measured by {\it Kepler}\ and the California group's respective surveys.
\end{abstract}

>>>>

This work aims to clarify the biases that stellar binarity introduces
to occurrence rates inferred from transit surveys.
In general, stellar multiplicity leads to diluted planetary 
radii, overestimated detection efficiencies, and an undercounted number of 
selected stars (but possibly an overcounted number of searched stars).
These effects skew occurrence rate measurements in different directions, and 
we develop simple models that allow us to understand the crucial effects.
For a model in which all stellar systems are either single or twin binaries, 
and all planets are identical, we find that ignoring binarity leads to an 
underestimate of the occurrence at the true radius by a multiplicative factor 
of $0.8$ (assuming a 10\% twin binary fraction).
Using more realistic models for the stellar population and planetary 
radii, we find that ignoring binarity leads to a $10$\,-\,$30\%$ 
underestimate of the number of planets per star, depending on the true radius 
distribution.
For the occurrence of Earth-sized planets, in our most realistic model the 
rate is underestimated by $\approx 15\%$~--~at present far smaller than 
systematic uncertainties on real $\eta_\oplus$ measurements.
For hot Jupiters, we find that the inferred occurrence rate is 
$\approx 1.3\times$ smaller than the true rate around single stars.
We suggest that this latter effect contributes to the discrepant hot Jupiter 
rates measured by {\it Kepler}\ and the California group's respective surveys.
\end{abstract}

### 2017-11-14 paper/model_2.tex
paper: improved clarity; corrected eqn error

@@

The rate density in this model, $\Gamma(r,q)$, is the sum of the rate 
densities for each system type:
\begin{equation}
\Gamma(r,q)
=
\delta(r_p) \times 
\frac{N_0 Z_0 + N_1 Z_1 + N_2 Z_2 p_2(q)}{N_{\rm tot}}
\label{eq:model2_rate_density}
\end{equation}
where $p_2(q)$ is expressed in terms of the mass ratio instead of the 
secondary star's radius for convenience ($q$ and $R_2$ are interchangeable).
The probability that a secondary hosts a planet, as a function of the mass 
ratio, is
\begin{align}
p_2(q) &= p({\rm has\ planet}\,|\,{\rm secondary\ with\ }q) \times
          p({\rm secondary\ with\ }q)
          \\
p_2(q) &\propto q^{\gamma + \beta} (1+q^\alpha)^{3/2}.
\label{eq:model_2_p_2}
\end{align}
We take first term, $p({\rm has\ planet}\,|\,{\rm secondary\ with\ }q)$, as a 
power law of $q$ with exponent $\gamma$.
The second term, $p({\rm secondary\ with\ }q)$, 
is the product of the volume limited probability, $q^\beta$, and a 
Malmquist-like bias, $(1+q^\alpha)^{3/2}$.
This is because the selected sample at a fixed planet radius and period is 
magnitude-limited.
We show the magnitude-limited mass ratio distribution for the $\beta=0$ case 
in Fig.~\ref{fig:q_distribn_mag_limited}.
We emphasize that in any Monte Carlo simulation of a transit survey, it is 
important to draw binaries from the correctly biased mass-ratio distribution 
(e.g., Bakos et al 2012, Sullivan et al 2015, Guenther et al 2017).

The occurrence rate corresponding to Eq.~\ref{eq:model2_rate_density}'s rate 
density for specific mass ratios of interest $q_{\rm min} < q < q_{\rm 
max}$ is
\begin{equation}
\Lambda|_{r_p,q_{\rm min},q_{\rm max}} = 
\frac{N_0 Z_0 + N_1 Z_1 + N_2 Z_2 f_2}
{N_{\rm tot}},
\end{equation}
for
\begin{equation}
f_2 \equiv
\int_{q_{\rm min}}^{q_{\rm max}} p_2(q) \,{\rm d}q,
\end{equation}
where $p_2(q)$ is given by Eq.~\ref{eq:model_2_p_2}, and must be normalized to 
unity.

\begin{figure}[!tb]
    \centering
    \includegraphics[width=0.6\textwidth]{figures/q_distribn_mag_limited.pdf}
    \caption{
        The distribution of the mass ratio for a magnitude limited sample of 
        binary stars. The underlying mass ratios are drawn from a uniform 
        distribution in a volume-limited sample, quite similar to that of 
        Raghavan 
        et al. (2010)'s Fig. 13.
        The entire bias can be understood analytically 
        (Eq.~\ref{eq:model_2_p_2},
        with $\gamma=0$).
        The numerical comparison uses a realistic empirical mass-luminosity 
        relation rather than $L=M^{3.5}$, found by fitting main sequence dwarf 
        mass luminosity data compiled from Torres et al. [2009] and Benedict 
        et 
        al. [2016].
    }
    \label{fig:q_distribn_mag_limited}
\end{figure}

\paragraph{What does the observers ignoring binarity infer?}
As a reminder, the observers compute an apparent rate density by correcting 
the number of detections per star for the transit probability and the 
detection efficiency.
In this process, the binarity-ignoring observers make the following errors:
\begin{enumerate}
\item They assume that they have selected $N_0+N_1$ stars, while they have 
actually selected $N_0+N_1+N_2$.
%
\item They assume that every star within the maximum selected distance is 
searchable.
This is true for single stars.
For binaries, the actual detection efficiency is the ratio of the number of 
searchable stars to the number of selected stars.
For primaries,
\begin{equation}
p_{\rm det,1} = \left(\frac{d_{\rm det,1}}{d_{\rm sel}}\right)^3
= \mathcal{D}_1^3 = (1+q^\alpha)^{-3},
\label{eq:model_2_p_det_1}
\end{equation}
where we have assumed that all single stars and primaries have the same 
properties, and have used $\ell = q^\alpha$.
For secondaries,
\begin{equation}
p_{\rm det,2} = \left(\frac{d_{\rm det,2}}{d_{\rm sel}}\right)^3
= \mathcal{D}_2^3 \left(\frac{R_1}{R_2}\right)^6
    \left(\frac{T_{\rm dur,2}}{T_{\rm dur,1}}\right)^{3/2} 
= (1+q^{-\alpha})^{-3} q^{-5},
\label{eq:model_2_p_det_2}
\end{equation}
where subscripts `1' and `2' correspond to primaries and secondaries, and we 
used the scaling relation $T_{\rm dur}\propto \rho_\star^{-1/3} \propto 
R_\star^{2/3}$.
%
\item They assume the transit probability for all stars to be $R_\star/a$. 
For planets orbiting secondaries at fixed period, the true transit probability 
is $q^{2/3} R_\star/a$, since the secondaries have smaller radii and masses.
%
\item The true planetary radii $r$ are interpreted as apparent radii $r_a$.
In binaries, the apparent radii depend on whether the host is the primary or 
secondary:
    \begin{align}
    r_a
    &=
    \left.
    \begin{cases}
    r_p (1+q^\alpha)^{-1/2} & \text{for } i=1,\ {\rm primary} \\
    r_p (1+q^{-\alpha})^{-1/2} q^{-1}, & \text{for } i=2,\ {\rm secondary}.
    \end{cases}
    \right.
    \label{eq:model2_ra}
    \end{align}
    The factor of $q^{-1}$ for the secondary case accounts for the observer 
    assuming all transit signals come from stars of fixed size.
\end{enumerate}

@@

Given a binary fraction, Eq.~\ref{eq:mu_model_2} fully specifies the 
``weights'' in Eq.~\ref{eq:model2_Gamma_a}.
The $I_1(r_a)$ and $I_2(r_a)$ terms are found by marginalizing over the joint 
distribution of apparent radius and mass ratio:
\begin{align}
I_i(r_a) &= 
\int_0^1 p({\rm has\ detected\ planet}, r_a, q | {\rm star\ is\ type\ }i)
    \,{\rm d}q,
\quad
{\rm for}\ i\in\{1, 2\}, \\
&=
\int_0^1 
    p({\rm has\ detected\ planet} | r_a, q, {\rm star\ is\ type\ }i) 
    \nonumber\\
    &\quad\quad\quad\quad\times p(r_a | q, {\rm star\ is\ type\ }i)
    \,p(q | {\rm star\ is\ type\ }i)
\,{\rm d}q.
\end{align}
The first term is the detection efficiency; the second is a $\delta$-function 
of the apparent radius; the last is the mass ratio distribution given by 
Eq.~\ref{eq:model_2_p_2}.
The analytic solution for $i=1$ is
\begin{equation}
I_1(r_a) = \frac{1}{\mathcal{N}_1} \left(\frac{r_p}{r_a}\right)^{-3}
\left( \left(\frac{r_p}{r_a}\right)^2 -1  
    \right)^{\frac{\gamma+\beta}{\alpha}},
\quad {\rm for}\ r_p/\sqrt{2} < r_a < r_p,
\end{equation}
where $\mathcal{N}_1$ is the normalization term of the binary mass ratio 
distribution (Eq.~\ref{eq:model_2_p_2}): $\mathcal{N}_1 = \int_0^1 
q^{\gamma+\beta} (1+q^\alpha)^{3/2} {\rm d}q$.

For $i=2$, there is no analytic solution, because evaluating the integral 
requires imposing the constraint that $r_a  = r_p 
(1+q^{-\alpha})^{-1/2}q^{-1}$. This equation can be re-written
\begin{equation}
\left(\frac{r_p}{r_a}\right)^2 = q^2 + q^{-\alpha + 2},
\end{equation}
which has no analytic solution for $q(r_a)$ except for special values of 
$\alpha$, the mass-luminosity exponent.
For $\alpha=3.5$, our nominal case, semianalytic solutions could be found.
However, since our main interest is in understanding the qualitative behavior 
of the solutions, we derive a few limiting cases, and then proceed 
numerically.

\paragraph{Limiting cases of rate density correction}
Recall that the rate density correction factor, $X_\Gamma$, is the ratio of 
the apparent to true rate densities.
We consider a ``nominal model'' in which the stellar population is similar to 
Sun-like stars in the local neighborhood:
${\rm BF}=0.44$, $\alpha=3.5$, $\beta=0$.
Our default assumption is also that the occurrence of planets is independent 
of stellar mass ($\gamma=0$), so secondaries have the same occurrence rate as 
primaries and single stars.
Under these assumptions, the true rate density is
\begin{equation}
\Gamma(r) \approx \delta(r_p) \left( Z_0 + Z_1 + 
Z_2 \right) / 3,
\label{eq:model2_Gamma_r}
\end{equation}
where the coefficients of $1/3$ are accurate to within one percent of the true 
coefficients.
Ignoring binarity, the observer finds an apparent rate density
\begin{equation}
\Gamma_a(r) = c_0 Z_0 \delta(r_p)
             +c_1 Z_1 I_1(r_a)
             +c_2 Z_2 I_2(r_a),
\label{eq:model2_Gamma_a_r}
\end{equation}
for $c_0\approx 0.49$, $c_1\approx 0.32$, and $c_2$ unknown.
%171011_integrals.nb
Evaluating the correction term at $r=r_p$, since $\lim_{q\rightarrow0} 
p_i(r_a)=0$ for $i\in\{1,2\}$,
\begin{equation}
X_\Gamma(r=r_p) \approx \frac{3c_0 \Lambda_0}{\Lambda_0+\Lambda_1+\Lambda_2}.
\end{equation}
If all the $Z_i$'s are equal, $X_\Gamma(r=r_p)\approx0.49$.
If there are no planets around the secondaries, $X_\Gamma(r=r_p)\approx0.74$.

@@

Once the free parameters are specified, the population of selected stars is 
constructed.
Each selected star is assigned a type (single, primary, secondary), a 
binary mass ratio (if it is not single), and the property of whether it is 
``searchable''.
The relative number of binaries to primaries is set by 
Eq.~\ref{eq:mu_model_2}.
The mass ratios are drawn from the appropriate magnitude-limited distribution 
(Eq.~\ref{eq:model_2_p_2}).
If it is single, a selected star is assumed to be searchable.
If it is a primary or secondary, its searchability is determined by a uniform 
draw from either Eq.~\ref{eq:model_2_p_det_1} or Eq.~\ref{eq:model_2_p_det_2}, 
as appropriate.

@@

\paragraph{Numerical results for Model \#2}
We first validate our model by ensuring it produces the analytically-predicted 
results for Model \#1, and the limiting cases of Model \#2 described above.
Following validation, assuming ${\rm BF}=0.44$, $\alpha=3.5$, 
$\beta=\gamma=0$, and taking $Z_0=Z_1=Z_2$, we compute the true and apparent 
occurrence rates over bins in planet radius.
The results are shown in Figs.~\ref{fig:errcases_model_2_linear} 
and~\ref{fig:errcases_model_2_log}.
Evidently, dilution produces a spectrum of apparent planetary radii. This 
leads to overestimated rates everywhere except where there 
are actually planets, where the rate is underestimated by a factor of two.

>>>>

\begin{figure}[!tb]
    \centering
    \includegraphics[width=0.6\textwidth]{figures/q_distribn_mag_limited.pdf}
    \caption{
        The distribution of the mass ratio for a magnitude limited sample of 
        binary stars. The underlying mass ratios are drawn from a uniform 
        distribution in a volume-limited sample, quite similar to that of 
        Raghavan 
        et al. (2010)'s Fig. 13.
        The entire bias can be understood analytically 
        (Eq.~\ref{eq:Malmquist_bias}).
        {\bf todo: make y axis label p(secondary with q).}
    }
    \label{fig:q_distribn_mag_limited}
\end{figure}

@@

The rate density in this model, $\Gamma(r,q)$, is the sum of the rate 
densities for each system type:
\begin{equation}
\Gamma(r,q)
=
\delta(r_p) \times 
\frac{N_0 Z_0 + N_1 Z_1 + N_2 Z_2 p_2(q)}{N_{\rm tot}}
\label{eq:model2_rate_density}
\end{equation}
where $p_2(q)$ is mass-ratio dependent shape function in secondaries. A fully 
general approach would parametrize $p_2(q)$ as a power law, $p_2(q) \propto 
q^\gamma$. For simplicity we will always take $\gamma=0$, the uniform case.
%We have written $\Gamma$ in terms of the mass ratio instead of the 
%secondary star's radius for convenience ($q$ and $R_2$ are interchangeable).
The interpretation of $Z_i$ is still ``the number of planets per star of 
type $i$'', but now for secondaries this must include a marginalization 
over both the mass ratio and the planet radius.

There is also a Malmquist bias in Eq.~\ref{eq:model2_rate_density}, hidden 
away in the numbers of selected stars.
This is because the selected sample at a fixed planet radius and period is 
magnitude-limited.
Given a binary, the probability of drawing a mass ratio $q$ in a 
magnitude-limited sample scales as
\begin{equation}
p({\rm draw\ }q | {\rm system\ is\ binary}) \propto q^\beta (1+q^\alpha)^{3/2}
\label{eq:Malmquist_bias}
\end{equation}
where $q^\beta$ is the volume-limited probability of drawing a binary of mass 
ratio $q$, and the latter term is the Malmquist bias.
We show the magnitude-limited mass ratio distribution for the $\beta=0$ case 
in Fig.~\ref{fig:q_distribn_mag_limited}.
We emphasize that in Monte Carlo simulations of transit surveys, it is 
important to draw binaries from the correctly biased mass-ratio distribution 
(e.g., Bakos et al 2012, Sullivan et al 2015, Guenther et al 2017).

The occurrence rate corresponding to Eq.~\ref{eq:model2_rate_density}'s rate 
density for specific mass ratios of interest $q_{\rm min} < q < q_{\rm 
max}$ is then
\begin{equation}
\Lambda|_{r_p,q_{\rm min},q_{\rm max}} = 
\frac{N_0 Z_0 + N_1 Z_1 + N_2 Z_2 f_2}
{N_{\rm tot}},
\end{equation}
for
\begin{equation}
f_2 \equiv
\int_{q_{\rm min}}^{q_{\rm max}} p_2(q) \,{\rm d}q,
\end{equation}
where $p_2(q)\propto q^\gamma$, and must be normalized to unity.

\paragraph{What do the observers ignoring binarity infer?}
As a reminder, the binary-ignoring observers compute an apparent rate density 
by correcting the number of detections per star for the transit probability 
and the detection efficiency.
In this process, they make the following errors:
\begin{enumerate}
\item They assume that they have selected $N_0+N_1$ stars, while they have 
actually selected $N_0+N_1+N_2$.
%
\item They assume that every star within the maximum selected distance is 
searchable.
This is true for single stars.
For binaries, the actual detection efficiency is the ratio of the number of 
searchable stars to the number of selected stars.
For primaries,
\begin{equation}
p_{\rm det,1} = \left(\frac{d_{\rm det,1}}{d_{\rm sel}}\right)^3
= \mathcal{D}_1^3 = (1+q^\alpha)^{-3},
\label{eq:model_2_p_det_1}
\end{equation}
where we have assumed that all single stars and primaries have the same 
properties, and have used $\ell = q^\alpha$.
For secondaries,
\begin{equation}
p_{\rm det,2} = \left(\frac{d_{\rm det,2}}{d_{\rm sel}}\right)^3
= \mathcal{D}_2^3 \left(\frac{R_1}{R_2}\right)^6
    \left(\frac{T_{\rm dur,2}}{T_{\rm dur,1}}\right)^{3/2} 
= (1+q^{-\alpha})^{-3} q^{-5},
\label{eq:model_2_p_det_2}
\end{equation}
where subscripts `1' and `2' correspond to primaries and secondaries, and we 
used the scaling relation $T_{\rm dur}\propto \rho_\star^{-1/3} \propto 
R_\star^{2/3}$.
%
\item They assume a transit probability for all stars of $p_{\rm tra} = 
R_\star/a$. 
For planets orbiting secondaries at fixed period, the true transit probability 
is $q^{2/3} p_{\rm tra}$, since the secondaries have smaller radii and masses.
%
\item The true planetary radii $r$ are interpreted as apparent radii $r_a$.
In binaries, the apparent radii depend on whether the host is the primary or 
secondary:
    \begin{align}
    r_a
    &=
    \left.
    \begin{cases}
    r_p (1+q^\alpha)^{-1/2} & \text{for } i=1,\ {\rm primary} \\
    r_p (1+q^{-\alpha})^{-1/2} q^{-1}, & \text{for } i=2,\ {\rm secondary}.
    \end{cases}
    \right.
    \label{eq:model2_ra}
    \end{align}
    The factor of $q^{-1}$ for the secondary case accounts for the observer 
    assuming all transit signals come from stars of fixed size.
\end{enumerate}

@@

Given a binary fraction, Eq.~\ref{eq:mu_model_2} fully specifies the 
``weights'' in Eq.~\ref{eq:model2_Gamma_a}.
The $I_1(r_a)$ and $I_2(r_a)$ terms are found by marginalizing over the joint 
distribution of apparent radius and mass ratio:
\begin{align}
I_i(r_a) &= 
\int_0^1 p({\rm has\ detected\ planet}, r_a, q | {\rm star\ is\ type\ }i)
    \,{\rm d}q,
\quad
{\rm for}\ i\in\{1, 2\}, \\
&=
\int_0^1 
    p({\rm has\ detected\ planet} | r_a, q, {\rm star\ is\ type\ }i) 
    \nonumber\\
    &\quad\quad\quad\quad\times p(r_a | q, {\rm star\ is\ type\ }i)
    \,p(q | {\rm star\ is\ type\ }i)
\,{\rm d}q.
\end{align}
The first term is the detection efficiency; the second is a $\delta$-function 
of the apparent radius; the last is the mass ratio distribution given by 
Eq.~\ref{eq:Malmquist_bias}.
An analytic solution can be found for $i=1$.
For $i=2$ there is no analytic solution, because evaluating the integral 
requires imposing the constraint that $r_a  = r_p 
(1+q^{-\alpha})^{-1/2}q^{-1}$. This equation can be re-written
\begin{equation}
\left(\frac{r_p}{r_a}\right)^2 = q^2 + q^{-\alpha + 2},
\end{equation}
which has no analytic solution for $q(r_a)$ except for special values of 
$\alpha$, the mass-luminosity exponent.
For $\alpha=3.5$, our nominal case, semianalytic solutions do exist.
However, since our main interest is in understanding the qualitative behavior 
of the solutions, we derive a limiting case, and then proceed 
numerically.

\begin{comment}
The analytic solution for $i=1$ is something like
\begin{equation}
I_1(r_a) = \frac{1}{\mathcal{N}_1} \left(\frac{r_p}{r_a}\right)^{-3}
\left( \left(\frac{r_p}{r_a}\right)^2 -1  
    \right)^{\frac{\gamma+\beta}{\alpha}},
\quad {\rm for}\ r_p/\sqrt{2} < r_a < r_p,
\end{equation}
where $\mathcal{N}_1$ is the normalization term of the binary mass ratio 
distribution (Eq.~\ref{eq:model_2_p_2}): $\mathcal{N}_1 = \int_0^1 
q^{\gamma+\beta} (1+q^\alpha)^{3/2} {\rm d}q$.
\end{comment}

\paragraph{Limiting case of rate density correction}
Recall that the rate density correction factor, $X_\Gamma$, is the ratio of 
the apparent to true rate densities.
We consider a ``nominal model'' in which the stellar population is similar to 
Sun-like stars in the local neighborhood:
${\rm BF}=0.44$, $\alpha=3.5$, $\beta=0$.
Our default assumption is also that the occurrence of planets is independent 
of stellar mass ($\gamma=0$), so secondaries have the same occurrence rate as 
primaries and single stars.
Under these assumptions, the true rate density is
\begin{equation}
\Gamma(r) \approx \delta(r_p) \left( Z_0 + Z_1 + 
Z_2 \right) / 3,
\label{eq:model2_Gamma_r}
\end{equation}
where the coefficients of $1/3$ are accurate to within one percent of the true 
coefficients.
Ignoring binarity, the observer finds an apparent rate density
\begin{equation}
\Gamma_a(r) = c_0 Z_0 \delta(r_p)
             +c_1 Z_1 I_1(r_a)
             +c_2 Z_2 I_2(r_a),
\label{eq:model2_Gamma_a_r}
\end{equation}
for $c_0\approx 0.49$, and the coefficients $c_1, c_2$ unknown.
%171011_integrals.nb $c_1\approx 0.32$
We can evaluate the correction term at $r=r_p$, since $\lim_{r_a\rightarrow 
r_p} I_i(r_a)=0$ for $i\in\{1,2\}$:
\begin{equation}
X_\Gamma(r=r_p) \approx \frac{3c_0 \Lambda_0}{\Lambda_0+\Lambda_1+\Lambda_2}.
\end{equation}
If all the $Z_i$'s are equal, $X_\Gamma(r=r_p)\approx0.49$.
If there are no planets around the secondaries, $X_\Gamma(r=r_p)\approx0.74$.

@@

Once the user specifies the free parameters, the program constructs a 
population of selected stars.
Each selected star is assigned a type (single, primary, secondary), a 
binary mass ratio (if it is not single), and the property of whether it is 
``searchable''.
The relative number of binaries to primaries is set by 
Eq.~\ref{eq:mu_model_2}.
The mass ratios are drawn from the appropriate magnitude-limited distribution 
(Eq.~\ref{eq:Malmquist_bias}).
If it is single, a selected star is assumed to be searchable.
If it is a primary or secondary, its searchability is determined by a uniform 
draw from either Eq.~\ref{eq:model_2_p_det_1} or Eq.~\ref{eq:model_2_p_det_2}, 
as appropriate.

@@

\paragraph{Numerical results for Model \#2}
We first validate our model by ensuring it produces the analytically-predicted 
results for Model \#1, and the limiting case of Model \#2 described above.
Following validation, assuming ${\rm BF}=0.44$, $\alpha=3.5$, 
$\beta=\gamma=0$, and taking $Z_0=Z_1=Z_2=0.5$, we compute the true and 
apparent occurrence rates over bins in planet radius.
The results are shown in Figs.~\ref{fig:errcases_model_2_linear} 
and~\ref{fig:errcases_model_2_log}.
Evidently, dilution produces a spectrum of apparent planetary radii. This 
leads to overestimated rates everywhere except where there 
are actually planets, where the rate is underestimated by a factor of two.
